### 2018-01-11 paper/meetingNotes.tex

%%%%% Meeting notes %%%%%

Jan 11, 2018

Discussion topics:

- What we are doing: local perturbation
- are we building a saliant map in our mind about the sentence?
  (which part of the sentence is most sensitive to the prediction)
  Should we just build it directly
- What information is important from Tao's point of view
- Comparison to other local explaianation techniques
- Where does the interpretability come from
    - It is interpretable because the application (model is still black box)
    - It is interpretable due to the inherent complexity of
      the model, sparsity linear model, rule list, dicision tree, etc.