### 2018-01-29 paper/intro_entailVis.tex

\firstsection{Introduction}
\maketitle

NLP application using attention
\begin{itemize}
    \item Neural translation
    \item Text summarization
    \item Text entailment
    \item Text comprehension (question answering)
\end{itemize

Attention type:
\begin{itemize}
    \item soft vs. hard
    \item local vs. global
    \item hierarchical
    \item attention over attention
    \item gated
\end{itemize}

How to interpret classification task
\begin{itemize}
    \item What feature / dimension / part-of-the-input should change to
    alter the outcome
    \item What Examples can you provide that is similar to the current
    instance

\end{itemize}