### 2018-01-29 paper/relatedWorks_entailVis.tex

\section{Related Works}

\begin{itemize}
    \item Depth learning vis
    \item Vis
\end{itemize}

Related works
\begin{itemize}
    \item Interacting with Predictions: Visual Inspection of Black-box Machine Learning Models,
    Use partial dependency (varying along one feature and observer how prediction changes)
    \item “Why Should I Trust You?” Explaining the Predictions of Any Classifier (LIME)

\end{itemize}