### 2018-02-05 paper/intro_entailVis.tex
add dependency tree to attenSen, but the parse tree compuation is duplicated

\firstsection{Introduction}
\maketitle

NLP application using attention
\begin{itemize}
    \item Neural translation
    \item Text summarization
    \item Text entailment
    \item Text comprehension (question answering)
\end{itemize

Attention type:
\begin{itemize}
    \item soft vs. hard
    \item local vs. global
    \item hierarchical
    \item attention over attention
    \item gated
\end{itemize}