### 2018-02-16 paper/intro_entailVis.tex
update paper layout

\firstsection{Introduction}
\maketitle

% NLP application using attention
% \begin{itemize}
%     \item Neural translation
%     \item Text summarization
%     \item Text entailment
%     \item Text comprehension (question answering)
% \end{itemize
%
%
% Attention type:
% \begin{itemize}
%     \item soft vs. hard
%     \item local vs. global
%     \item hierarchical
%     \item attention over attention
%     \item gated
% \end{itemize}