### 2018-03-03 paper/perturbation.tex
fix resizing

\section{Analysis of the Model Via Perturbation}

%%%%%%%% Overview figure for the pipeline %%%%%%%%%%
% \begin{figure}
%     \includegraphics{/path/to/figure}
%     \caption{}
%     \label{}
% \end{figure}

\subsection{Input Perturbation}
\begin{itemize}
    \item Just looking at the failure cases is not enough, the prediction maybe correct for the wrong reason
    \item sensitivity of the model with respect of
\end{itemize}

Iterate the model design and debug the system hinged on the ability to quickly and identify the errors made by a model.

Perturbation the input is what NLP researchers have subconsciously been doing to study and test a model.

Interpret/probe the relationship between attention and the prediction result

\subsection{Attention Perturbation}

\subsection{Prediction Perturbation}