### 2018-03-23 paper/intro.tex

\firstsection{Introduction}
\maketitle

% (What is the high level motivation)
% What is the obsatcle for effectively apply machine learning model
With the recent advances in neural network based model, machine learning has gained unprecedented popularity, and its adaption permeated many fields of studies.
%
However, from researchers to practitioners, one often need to overcome many obstacles during the training, debugging, and tuning processes to unleash the full potential of these models.
%
Interpreting the internal mechanism and analyzing how predictions are made are critical to both the design and deployment of a model.
More importantly, the ability to pinpoint where and how an error is made and come up with hypotheses for the cause of failure is the key to identify the limitation and improve upon existing models.
However, providing meaningful answers to these questions are incredibly challenging and often regarded as impossible by many.

Recently, sizable research has been developed to combat the model interpretability challenges. Most notably, the works~\cite{SimonyanVedaldiZisserman2013, ZeilerFergus2014, YosinskiCluneNguyen2015, OlahMordvintsevSchubert2017, LiuShiLi2017, OlahSatyanarayanJohnson2018, BilalJourablooYe2018} on interpreting convolution neural network (CNN) model from both machine learning and visualization community have revealed the enormous potential and immediate benefits of model interpretation. At the same time, these works also remind us how little we truly understand about the inner mechanism of these deep neural networks.

Compare to image classification task, natural language processing (NLP) systems have innate restrictions (e.g., the discrete nature of words) that makes the road to interpretation more treacherous. For example, \emph{feature visualization}~\cite{OlahMordvintsevSchubert2017}, an often deployed technique for illustrating what image features a given part of the network (e.g., neuron, layer, or channel) capture, can not be readily generalized to natural language. Image (pixel values) represents a continuous solution space, in which human can quickly recognize the existence (or the absence) of patterns or shapes. For natural language, even though we can encode a word as a vector~\cite{MikolovSutskeverChen2013, PenningtonSocherManning2014}, the word embedding space is still discretely defined, in which the interpolation of two vectors (i.e., two words) do not hold clear meaning.
%
These restrictions call for new avenues for solving the interpretation challenges of natural language models, which motivate the proposed work.

% Beside explore the internal states of the model, understand how different parts of the model work together and how one may affect one another is as or if not more important.
Moreover, despite many recent advances, the majority of the existing techniques study the model as an invariant object, where the model's behaviors are recorded and analyzed in an offline fashion.
%
However, the exploratory nature of the model interpretation analysis often leads to many \emph{``what if ...''} type of questions. Such as, what if we perturbed the current input? Will the prediction be stable? What if we change one of the critical internal states of the model, how would the modification affect the prediction? What if the current prediction is wrong, how and where could we apply minimal change to the model to produce the correct result? And how would it affect the internal state we care about? These type of queries form a natural way to gain understanding and come up with hypotheses of the model mechanisms by \emph{interrogating} how the components of a model interact with each other in a dynamic setting. Hypothetically, we can code up specialized experiments for each of these scenarios. However, such a process is not only tedious but also ignores the iterative nature of the exploratory analysis. Often, these questions are not pre-determined. Instead, new exploration paths arise as we investigate and analyze previous observations.

In this work, we aim to provide immediate and informative answers to these \emph{``what if''} questions by combining the expressive power of visualization and direct online query/optimization of the neural network model. Instead of viewing the model as an invariant object, we approach the interpretability challenge by studying them in a dynamic environment. By employing a perturbation-driven paradigm, we probe the internal states of the model and examine how changes in one part of the pipeline (the input, internal states, and output prediction, see Fig.~\ref{fig:modelPipeline}) affect others, which in turn provides a new perspective to address the model interpretation problem.

We implement our vision of the perturbation-driven exploration in an interactive visualization system for natural language inference models~\cite{Parikh2016}.
However, the components of the visualization and the overall paradigm can be readily extended to other NLP tasks, such as neural machine translation, and text summarization, etc.
%
In the simplest form, the inference task ask the question about whether if we can infer the content of \textbf{sentence B} from \textbf{sentence A} (\emph{entail}), or are they contradict with each other (\emph{contradiction}), or are they talk about different/unrelated things (\emph{neutral}). Language inference addresses the fundamental challenge of measuring sentences' semantic similarity, and is one the core natural language processing tasks (see detail in Section~\ref{sec:languageInference}).

One recent advance in neural natural language process models is the introduction of attention mechanisms~\cite{VaswaniShazeerParmar2017} (Section~\ref{sec:attention}). Intuitively, the \emph{attention} help capture which part of the input the model deemed more important for making the right prediction. The attention is often represented as weights for individual words or pairs of words (i.e., the alignment between words in different sentences).
%
%\shusen{seems repeated some of the previous points}
There have been many speculations on how attention works in various models. By utilizing the proposed tool, the domain experts study how changes in sentence input, attention, or prediction affect each other, which help them develop deeper intuitive and alternative hypotheses. 
%we can perturb the attention and observer the change in the prediction, perturb the sentence input and observe the effects on the attention. And more importantly, we also allow the user to change prediction, where we apply a constrained optimization to find the least amount of parameter update to produce the user assign label. 
%Another word, 
%Such type of exploration is not only essential to interpret the language inference problem, but can also be generalized to other NLP model utilized the attention mechanisms.
Moreover, we enhanced the standard visual encoding (e.g., as a bipartite graph or as a matrix) of the attention matrix by overlaying sentence linguistic structure to allow grammar guided simplification of the visual representation.
%
Finally, as discussed in Section~\ref{sec:caseStudy}, the ability to examine how attention corresponds to the grammatical structure also enables domain experts to speculate the potential benefit of including the linguistic structure in the design of attention component of the model.

In summary, the key contributions of the proposed works is the following:
\begin{itemize}
  % \item Distill the visualization tasks for making sense of the role of attention in natural language inference (section~\ref{sec:task});
    \item Introduce the perturbation-driven exploration paradigm for exploring natural language processing models;

    \item Perform error analysis by examining stability of the predictions under perturbation;

    \item Enhance the visual encoding of the attention by overlaying sentence linguistic structure to allow grammar guided simplification;

    \item Enable the interrogation of the relationship between input sentences, model internals, and the prediction by interactively exploring how the change in one place affect other parts of the model.
\end{itemize}
%
% In addition,

%\begin{itemize}
%    \item Study the classifier  vs. single instance evaluation
%    \item Model agnostic vs. reveal internal states
%    \item View model as invariant object vs. interrogate the model via partial/full updates
%\end{itemize}

%Questions such as, how stable predictions are concerning small perturbation in the input, how are these predictions arrived (i.e., the model may come up with the correct prediction by pickup unintended feature/alignment), and what are the often made mistakes, are all essential in systematic examination and evaluation of the model.
%%%%%%%%% accessing internal states are important %%%%%%%%%%%%

%%%%%%%%%% interrogate relationship between different components, modified different components %%%%%%%%%%%

% ######## Why use perturbation ###########
% Iterate the model design and debug the system hinged on the ability to quickly
% identify the errors made by a model.
%
% Perturbe the input is what NLP researchers have subconsciously been doing to study and test a model.
%
% Interpret/probe the relationship between attention and the prediction result

% On the one hand, recognize where the model fail is non-trivial. The standard evaluation approach (i.e., the prediction accuracy on the test set) provides limited information and cannot answer many important questions (e.g., how stable the prediction is, which part of the model may contribute to the failure) that are essential to the model diagnostic process. On the other hand, infer the cause of model failure is even more challenging. Neural network models are often referred to as black boxes due to the hard to recognize  difficult to interpret parameter, hypothesizing what happened in prediction process require the ability to peek inside the model and understand the intricate relationships among model inputs, critical internal mechanisms, as well the output predictions.