### 2018-05-30 EMNLP-demo-paper/intro.tex
update intro
feedback from chris

\section{Introduction}

Deep neural networks have been successfully applied to various natural language processing tasks,
%where end-to-end training become a primary option.
and as the design of neural networks evolves, there is a clear trend of increasing model complexity in terms of both the architecture and the parameters count. 
%
The expressiveness of these models helps improve the prediction performance, there is a fundamental lack of interpretability leads many researchers to consider neural network models black boxes. 
%
Despite the difficulties, interpreting the model internals and reason about predictions are essential for understanding the limitation of the models and improving upon the existing designs.

%Without understanding the background meaning that features carry, visualization relies
%on proper quantification analysis (cite). Such analysis is not easy to come up with,
%and is tied to specific task (cite).

Recently, the attention networks have become a widely-adopted mechanism \cite{bahdanau2014neural,seo2016bidirectional,Parikh2016, VaswaniShazeerParmar2017}. The attention not only improves the model performance but also yields interpretable intermediate representations (i.e., alignment among words). As a result, attention is often considered as a natural interface for analyzing the internals of neural networks. Conduct analysis directly on the raw attention values can be challenging for human users, and therefore, visual representations that highlight word alignments between sentences have been proposed, such as bipartite graph and heatmap of the attention matrix \cite{LiChenHovy2015, li2016understanding, lee2017interactive}.  However, many challenges remain. Firstly, for some NLP tasks, the word sequences pair for which the attention is computed can be highly asymmetrical (i.e., in machine comprehension, the context paragraph can be much longer than the question sentence), which the standard visual encoding cannot adequately handle.  Also, many previous works present the attention in a static setting. However, the ability to provide an interactive environment, in which the user can instantaneously look at how changes in the input affect the attention, and how small variations in the attention alter the prediction, is crucial for interpreting the model. Finally, many previous visualization efforts, despite revealing many exciting results, focus more on illustrating potentially useful visualization technique than on providing a flexible software tool that can be easily integrated into an existing code base. As a result, applying these proposed techniques to real-world examples may involve substantial engineering effort that could prevent wide adoption.

To address these challenges in interpreting attention-based models, 
we introduce a flexible Python library that allows the user to easily generate a web-based interactive visual analytics environment for various usage via a composable visualization component design.
%
To facilitate exploratory analysis (see Figure~\ref{fig:modelPipeline}), we employ a fully interactive pipeline, in which the user can perturb the input text (via two perturbation schemes discussed in Section~\ref{sec:perturb}) and observe the effects on the attention and prediction, or modify the attention to understand how it affects predictions.
%
Finally, we demonstrate our system on two primary NLP tasks: natural language
inference (NLI) and machine comprehension (MC). 
%For the NLI task, we visualize
%decomposable attention networks which represents a series of strongly performing models \cite{Parikh2016}.
%For the MC task, we present your visualization methods on bidirectional attention flow model,
%which also represents a line of state-of-the-art models \cite{Seo2016}.

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{pipeline}
 \vspace{-3mm}
 \caption{
 Perturbation-driven interrogation of the end-to-end models that follow the encoder, attention, and classifier structure. The user can generate a small perturbation of the input (i.e., replacing synonyms, paraphrasing), edit the attention, and observe the changes.
 }
 \vspace{-3mm}
\label{fig:modelPipeline}
\end{figure}

%In summary, our key contributions are:
%\begin{itemize}
%	\item Introduce an flexible visualization system for creating customized interactive visual analytic environment for attention centric neural networks models;
%	\item Illustrate the importance of envire
%	\item Demonstrate our system on two major NLP tasks with strong-performing
%	models. We observe functional limitations of these models which can be helpful
%	for future improvement on modeling.
%\end{itemize}