### 2018-05-31 EMNLP-demo-paper/result.tex
update result

%\begin{figure*}[t]
%\centering
%\vspace{-2mm}
% \includegraphics[width=1.0\linewidth]{NLI_MC_interface}
%  \vspace{-6mm}
% \caption{
%Illustration of different configurations for the natural language inference and machine comprehension tasks.
%}
%\label{fig:pipelineUpdate}
%\end{figure*}
\begin{figure*}[t]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{NLIexample}
  \vspace{-7mm}
 \caption{
An illustration of the attention editing process.
The dependency structure is shown in (a), where the two ``green'' are decorating different nouns.
By removing the ``wrong'' alignment in (b), the originally prediction \emph{entailment} is corrected to \emph{netural} in (c).
}
\vspace{-2mm}
\label{fig:NLIexample}
\end{figure*}

\section{Applications}
We demonstrate the proposed visualization system on the decomposable attention network~\cite{parikh2016emnlp}
for the NLI task and the bi-directional attention flow network~\cite{Seo2016} for the MC task.

\subsection{Natural Language Inference}
\label{sec:NLIexample}
The NLI task predicts the relationship between a premise sentence (\textbf{P}) and a hypothesis sentence (\textbf{H}).
The attention matrix captures the alignment of words between these two sentences.
Here we give an example of how a wrong prediction can be correct by editing the attention value.

%The model produces two attention matrices, representing alignment from premise to hypothesis, and from hypothesis to premise.
%as either \emph{entailment}, \emph{contradiction}, or \emph{neutral}.
%The model can be formalized as the following:
%%\begin{align}
%	$P^\prime, H^\prime = f(P), f(H)$,
%	$\overleftarrow{A} = P^\prime \cdot H^\prime$,
%	$\overrightarrow{A} = H^\prime \cdot P^\prime$,
%	$y = g(P, H, P^\prime, H^\prime, \overleftarrow{A}, \overrightarrow{A})$,
%%\end{align}
%where $P$ and $H$ are input embedding matrices for the premise and the hypothesis, $\overleftarrow{A}$
%and $\overrightarrow{A}$ are attentions, and $y$ is the predicted probabilities for candidate classes.
%Our system visualized the bidirectional attentions and their interaction with output distribution over labels.

As illustrated in Fig.~\ref{fig:NLIexample}(a), the input sentence pairs (\textbf{P}:``A woman in a green jacket is drinking tea.'' \textbf{H}:``A woman is drinking green tea.'') is predicted to be entailment, which is incorrect.
By examining the attention, we can see the word \emph{green} in ``\emph{green} jacket" is aligned to the \emph{green} in ``\emph{green} tea". However, these two ``green'' decorate different nouns, which lead to the wrong prediction. The grammar structure is captured by the dependency tree, i.e., \emph{green} in \textbf{P} is attached to ``jacket''. However, the model does not has access to the grammar information and mistakenly assume the two \emph{greens} are describing the same thing (thus predict \emph{entailment}). 
%
To correct the mistake, we can edit the attention value and remove the align between these two "green" (see (c)(b)).
As expected, the prediction label is corrected (\emph{neutral}).

\subsection{Machine Comprehension}
\label{sec:MCexample}
In machine comprehension task, two sequences of texts are given: context and question.
The goal is to select a span of text from the context that answers the question. 
%We provide a simple model formulation and refer the reader to the paper for details.
%%\begin{align}
%	$C^\prime, Q^\prime = BiLSTM(C), BiLSTM(Q)$,
%	$\overleftarrow{A} = u(C^\prime, Q^\prime)$,
%	$\overrightarrow{A} = v(C^\prime, Q^\prime)$,
%	$s = m(C^\prime, Q^\prime, \overleftarrow{A}, \overrightarrow{A})$,
%	$e = n(s, C^\prime, Q^\prime, \overleftarrow{A}, \overrightarrow{A})$,
%%\end{align}
%where $C$ and $Q$ are embedding matrices for the context and the question,
%$\overleftarrow{A}$ and $\overrightarrow{A}$ are bidirectional attention flows,
%$s$ and $e$ are probabilities for starting and ending indices of the answer span.
%Our system reveals the internal states of $\overleftarrow{A}$, $\overrightarrow{A}$,
%$s$ and $e$.
The attention information is encoded as a bi-directional alignment (i.e., from context to question and vice versa). 
%The key difference is how the question to context alignment is computed (see \cite{Seo2016} for more details).
Here, we refer the context to question attention as \emph{att1}, and question to context attention as \emph{att2} (the \emph{att2} is a vector instead of a matrix, see \cite{Seo2016} for more details. In this demo, we applied a min max normalization for \emph{att2} after the softmax layer to better distinguish different attention values).

\begin{figure*}[t]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{MC_interface}
  \vspace{-7mm}
 \caption{
In the machine comprehension visualization interface (a), the $p1$, $p2$ colored bar (in $a3$) illustrate the predicted start and end index of the answer in the context (the deeper the red the higher the probability). The most likely answer is shown in ($a1$). The global and local attention are visualized by ($a2$, $a3$).
%
We can evaluate the robustness of the prediction by perturbing the question sentence ($b1$, $b3$). As illustrated in ($b1$, $b2$) by removing the word ``partial'' the model still find the correct answer (albeit different, as the sentence perturbation changes the exact meaning of the question). 
}
\label{fig:MCexample}
\end{figure*}

As illustrated in Figure~\ref{fig:MCexample}, we represent \emph{att2} as colored bars with \emph{yellow-green-blue} colormap. Each rectangular bar corresponds to one sentence. The user can focus on individual sentences by click on the bar (see Figure~\ref{fig:MCexample}).
%The proposed visualization help reveal the potential alignment issues in the machine comprehension model.
The $p1$, $p2$ colored bars (\emph{white-red} colormap) illustrate the predicted beginning and ending index for the answer (deeper the red the higher the probability). %The span of text defined by the beginning and end index with the highest probability is the answer.
%
As we can see in Figure~\ref{fig:MCexample} ($a3$), there are two spans of text that correspond to higher probabilities. 
In (a3), we examine the sentence that contains the answer, which exhibits excellent alignment with the question (e.g., ``Two'' is aligned with ``many''). 
%The spans of text with high probability are ``Two'' and ``No'' (highlight in orange), both align well with the word ``many''.
An interesting observation is that the number ``9'' (in ``Falcon 9') is also aligned with ``many'', which may potentially lead to problems.

The user can explore the robustness of the model by examining how the prediction varies when the question is perturbed.
%
As illustrated in ($b1$, and $b2$), the perturbation removes the word ``partial'' in the original sentence, which leads the model to produce a different yet correct answer (``No''). Refer back to ($a1$), we can see the word ``No'' exhibits the second highest probability for the original question.
%
The user can also manually edit the text. As shown in ($b3$). The model still produces the correct answer when changing the question from ``how many'' to ``which''.
%One possible explanation is the word ``Two'' has much higher \emph{att2} attention value compare to ``9'', which may contribute to the current outcome.