### 2018-09-06 gateVis-paper/intro.tex
add template

The, recently several visualization work has focused on visualizing the recurrent neural network such as the long-short term memory (LSTM) network